---
title: "Math 218: Final Report"
subtitle: "Project Haydoja"
author: "Peyoja and Hayden"
output: pdf_document
date: "?"
---

```{r packages-data, include = F}

library(tidyverse)
dat <- read_csv("data/train.csv")

```

## Introduction

For our final project, we decided to look at the Ames Housing dataset, which includes extensive information on the sale of residential properties in Ames, Iowa from 2006 to 2010. It contains 1460 observations with 79 variables on various house specifications. The variables are a mix of continuous, categorical and discrete types. The data is publicly available on Kaggle - below is a link to the dataset itself which includes a description of all the data fields:

https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques/data

The goal of our project is to implement as many of the modeling techniques we've covered in this course as possible. To do so, we decided to focus on two modeling problems: one regression and one classification.

Below are our research questions:

1. Which modeling technique most accurately predicts housing prices using the other available housing features?

2. Which modeling technique most accurately classifies houses by neighborhood based on the other available housing features?

## EDA

```{r echo = F}
# Notice that you don't see this code chunk when you knit, but the code still executes! This is what the echo = F statement in the code chunk header does: it runs the code, but doesn't display it in the knitted document.
x <- 1+1
```

```{r Correlation}

```


### Data Cleaning

```{r}

#Convert 'Alley' to binary
dat$Alley <- ifelse(is.na(dat$Alley), 0, 1)

#Get rid of Utilities variable
dat %>%
  group_by(Utilities) %>%
  summarise(count = n())

dat <- select(dat, -Utilities)

#Dropping Condition2 variable
dat %>%
  group_by(Condition2) %>%
  summarise(count = n())
  
```


## Methodology

## Results

## Discussion