---
title: "Math 218: Project Proposal"
subtitle: "Team Haydoja"
author: "Payoja Adhikari and  Hayden Hunskor"
output: pdf_document
date: "11/13/2022"
---

```{r packages-data, message = F}
library(tidyverse)
library(dplyr)
library(tidyr)
```

### Data description

We picked the Ames Housing dataset compiled by Dean De Cock, which is publicly available on Kaggle. The dataset describes the sale of residential properties in Ames, Iowa from 2006 to 2010. It contains 1460 observatios with  79 variables on various house specifications. The variables are a mix of continuos, categorical and discrete types. Here is a link to the dataset.

https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques/data

### Define variables you want to explore
Some of the variables that we want to explore in our data sets are:

##### Categorical Variable:

- Street: Type of road access to property (Gravel/Paved)

- Utilities: Type of Utilities available

- Neighborhood: Physical locations within Ames city limits

- Heating: Type of heating

- CentralAir: Central air conditioning (Yes/No)

- PavedDrive: Paved driveway

- GarageFinish: Interior finish of the garage

- SaleType: Type of sale

- SaleCondition: Condition of sale

##### Continuous Variable:

- SalePrice: Price of house 

- PoolArea: Pool area in square feet

- Lot Area: Lot size in square feet

- TotalBsmtSF: Total square feet of basement area

- Garage Area :Size of garage in square feet

- GrLivArea: Above ground living area square feet

##### Discrete Variable


- FullBath: Number of full bathrooms above ground

- HalfBath: Number of half bathrooms above ground

- BedroomAbvGr: Number of bedrooms above ground

- KitchensAbvGr: Number of kitchens above ground

- FirePlaces: Number of fireplaces

- YrBuilt: Original construction date

- YrSold: Year house was sold


* We eventually want to run a lasso regression for which we will be utilizing all 79 variables.

### EDA
```{r}
train<-read_csv("data/train.csv")
#find variables that are numeric
numeric_var<-names(train)[which(sapply(train, is.numeric))]
#find variables that are categorical
categorical_var<- names(train)[-which(sapply(train, is.numeric))]
#numeric variable that we want to work with
var<- numeric_var[-c(1:3,5:6,8,10:12,16,18:19,27,35:36)]
#summary table for numeric variables. The table includes some variable not mentioned earlier
train %>%
  select(var)%>%
  summarise_all(list(mean = mean, std = sd), na.rm=TRUE)%>%
  pivot_longer(cols = everything(), 
               names_to = c('col', '.value'), 
               names_sep = '_')%>%
print(n=34)
```
### Correlation between housing prices and some of the discrete and categorical data
```{r}
library("ggcorrplot")
correlation<-train%>%
  select(SalePrice, PoolArea,LotArea, TotalBsmtSF, GarageArea,GrLivArea, FullBath,HalfBath,BedroomAbvGr,KitchenAbvGr,Fireplaces)
corr_mtx <- cor(correlation, use = "pairwise.complete.obs")
ggcorrplot(corr_mtx, # input the correlation matrix
           hc.order = TRUE, 
           type = "upper", 
           lab = TRUE,
           method = "circle", 
           colors = c("tomato2", "white", "springgreen3"))
```
As seen from above, housing prices seem to be the most correlated (positively) with above ground living area (in sq. ft), total basement area(in sq. ft) and garage area (in sq. ft). The price is also moderately correlated with number of full bathrooms.




### Potential Research Questions

How does housing prices differ in Ames, Iowa based on the different housing features?
